Latin epigraphy for the visually impaired. New technologies to favour universal accessibility

### 2016-01-19 source/44/Damnjanovic_et_al_EAGLE2016.tex
papers revised
minor fixes

%%%%%%%%
%TITLE:%
%%%%%%%%
\title{INTEGRATION OF MULTIMEDIA COLLECTIONS AND TOOLS FOR INTERACTION WITH DIGITAL CONTENT.
The case study of the Archaia Kypriaki Grammateia Digital Corpus}
\author[cyi]{Valentina Vassallo\corref{first}}
\author[cyi]{Uros Damnjanovic}
\author[cyi]{Sorin Hermon}
\date{2015-11-26}
\cortext[first]{Corresponding author. Email: (u.damnjanovic, v.vassallo, s.hermon)@cyi.ac.cy}
\begin{document}

\maketitle
\begin{abstract}
Supporting a discovery, use, and navigation of digital collections is a fundamental part of providing access and
encouraging inquiry, interpretation, and knowledge. In this paper we present our efforts to store and explore
multimedia collections of archaeological data. Particularly, the case study of the Archaia Kypriaki Grammateia
epigraphic collection is presented. Our work can be seen twofold. One aspect of our work is to provide a place where
the data coming from various sources can stored and accessed. Another aspect is to provide users with means to explore
this data. We argue that currently digital libraries are constrained by their webpage-based paradigm, thus not
providing the means for utilizing the full potential of the heritage data.
\end{abstract}
\keywords{Digital libraries, Data repository, Epigraphy, User interfaces, Data visualization, Interaction}

@@

\noindent Nowadays digital information format found its ways to all the spheres of our lives. Cultural heritage is no exception to
this. Gigabytes of data are being created on an every-day basis. For some time now, digital libraries are used to store
and provide access to digital heritage data. However, existing digital libraries often fail to provide the ability to
iteratively explore items, compare data trends, and engender the wisdom that comes from exploring data in new ways.
Current digital libraries are limited by their inadequate webpage-based paradigm, and it is easy for even the most
experienced scholar to get lost \citep{augmenting_2009}.
To overcome this issue, new generation of digital
libraries should emerge that will serve as an interactive knowledge environment in which knowledge is created by
interacting with the data. Digital collections often lack features for deeper quantitative and qualitative analysis,
and even very useful functions, such as the ability to annotate or bookmark content, are often not supported. After
digitisation, these collections are typically monolithic, difficult to navigate, and can contain text which is of
variable quality in terms of language, spelling, punctuation and consistency of terminology.

According to \citep{fast_interaction_2010} digital libraries are seen as a store of epistemic potential with interaction having
the role to reveal much of the hidden complexity. In \citep{buchel_making_2014} the authors demonstrated the role that a
set of interactions can play in supporting users' understanding of spaces and their formation of cognitive maps when
working with map-based visualisations. According to \citep{algee_viewshare_2012} interactive visualizations empower users to
discover meaning and patterns within digital collections using dynamic, interactive displays. One example of utilizing
the interactive graphics in a real world digital library is presented in \citep{hienert_integrating_2012}. Authors developed
interactive visualization tools that support exploration of search queries and search results and help users in
formulating new queries. Another example, the application called DaisyViz shown in \citep{ren_daisyviz:_2010}, tries to help
users acquire better insights of the data by enabling users to rapidly develop domain-specific information
visualizations without traditional programming. In \citep{nasar_conceptual_2011} authors proposed a conceptual framework that
uses interactive visualizations for managing personal collections of images and videos with focus on data re-finding
and improved filtering. INVISQUE \citep{wong_invisque:_2011} is a novel system designed for interactive information
exploration. Instead of a conventional list-style arrangement, in INVISQUE information is represented by a
two-dimensional spatial canvas, with each dimension representing user-defined semantics. Search results are presented
as index cards, ordered in both dimensions. Intuitive interactions are used to perform tasks such as keyword searching,
results browsing, categorizing, and linking to online resources such as Google and Twitter. An approach for
overview-first exploration of data collections based on user-selected metadata properties is presented in \citep{bernard_content-based_2012}. In a 2D layout representing entities of the selected property are laid out based on their similarity with
respect to the underlying data content. The display is enhanced by compact summarizations of underlying data elements,
and forms the basis for exploratory navigation of users in the data space. In \citep{cao_whisper:_2012} authors propose a
novel visual design termed ``Whisper'' to fulfil the need for tracing information diffusion processes in social media, in
a real time manner. BirdVis \citep{ferreira_birdvis:_2011} leverages visualization techniques and uses them in a novel way to
better assist users in the exploration of interdependencies among model parameters. Furthermore, the system allows for
comparative visualization through coordinated views, providing an intuitive interface to identify relevant correlations
and patterns.

@@

\noindent STARC repository was initially developed to store the data coming from the daily work of the research group and as
content aggregator for various digital libraries projects: data from archaeological excavations, museums artefacts,
epigraphic corpora, etc. Data stored in the repository ranges from high resolution images, 3D models, texts, maps,
videos and audio resulting from various data acquisition procedures such as photogrammetry, laser scanning, 3D
modeling, photography, sketching, drawing and so on. \ Every item in the repository is described using a metadata
schema that depends on the type of data and according to the needs of the project. For most of the items, we are using
our own metadata schema (STARC metadata schemas) \citep{ronzino_metadata_2012}; \citep{vassallo_revealing_2013}, that was developed to
thoroughly describe various aspects of data creation process. \ Once data is uploaded and available, it can also be
used in different systems. For example, the repository serves as source of data for aggregation procedures by which our
data feed the Europeana portal (\url{http://www.europeana.eu/portal/}) through different digital libraries projects (e.g.
Athena, CARARE, Linked Heritage, AthenaPlus, EAGLE). In order to facilitate the migration of data from heterogeneous
sources, the repository provides ingestion capabilities, that enable easy transfer of data from any available data
end-point. Another aspect of the back-end is the user management. It deals with providing different access levels for
different users’ groups. It also manages personal user space, where users can add their own information to the
repository and have access to it when exploring the data. Data annotations and personal collections are some of the
examples of how users can add information to the repository and use it for data exploration.

@@

\noindent We wanted to provide users with number of tools that they can use while exploring the repository. In traditional web
based system, navigating from one tool to another usually means that we click on a link and then move to another page
where the tool is available. Once the user is on a page s/he can start interacting with the tool by adding or modifying
data or by performing any action relevant to the selected functionality. The process of exploring the repository starts
by running the initial query (Fig.~\ref{fig:1}).

Once the results are returned and displayed, the exploration process can start. By default the map functionality is
available in the view. When a new tool is added to the screen, it automatically gets an access to all the data
available in the current search. It also gets notified on any other tools that might be active in the view, so the
necessary communication channels are established. Once the results are ready, there are number of filters available to
filter out the data, either by data type, collection. Search form can be used to filter out the data by typing the text
used to filter the data. As the user types, the data is automatically updated to show only those items that are related
to the text in the search field. The user can then run a new search repeatedly, and the search history functionality
will store all those searches for later reference. The map tool (Fig.~\ref{fig:1}) is used to show information about the
geo-referenced data. It provides two main functions. One is that it accompanies the search tool. Whenever the new
search is ran, and the results become available, the map tool updates itself with new data, showing the geographical distribution of the search results. The user can then use the map to filter out the data by clicking on the
markers on the map. Also the map can also be used as a search tool. By selecting the search map mode, the rectangle
mask appears on the map, that is used to define the region of interest on the map. This region is then used as an
additional parameter for the search.

\begin{figure}[!bp]
\centering
 \includegraphics[width=\columnwidth]{DamnjanovicetalEAGLE2016-img001.png}
\caption{The exploration process starts by running the search. Once the results are available the user can add or remove
exploration tools. By default map tool is present in the view. The map tool is used to mark the geo-referenced data on
the map, and also to perform geo-search by defining the region of interest. }
\label{fig:1}
\end{figure}

@@

\begin{figure}[!bp]
\centering
 \includegraphics[width=\columnwidth]{DamnjanovicetalEAGLE2016-img002.png}
\caption{The annotation tool. The user first selects the object in the search results, then selects the region of the
object and then adds the annotation. The annotation can be accessed when the annotated object is accessed, and by
specifying in the search that the search should also include annotations. }
\label{fig:2}
\end{figure}

@@

\begin{figure}[!bp]
\centering
 \includegraphics[width=\columnwidth]{DamnjanovicetalEAGLE2016-img003.png}
\caption{Visualization tool. Visualization tool is used to accompany the search tool by showing live visualizations as
the search results change. When the search is ran and the search results are available the visualization tools
automatically update itself and shows the visualisation related to the current search results. }
\label{fig:3}
\end{figure}

@@

\begin{figure}[!bp]
\centering
 \includegraphics[width=\columnwidth]{DamnjanovicetalEAGLE2016-img004.jpg}
\caption{Multimedia data constituting the AKG digital collection stored in the STARC repository.}
\label{fig:4}
\end{figure}

@@

The metadata schema is organized in groups corresponding to different research areas clustered in wrappers and
sub-wrappers, in order to fully describe all information that the ‘asset inscription’ contains. This schema,
distinguished by its multidisciplinary structure, include all disciplines relevant for description and representation
of epigraphies: archaeology (investigating for example the context of the finds), philology \ (analyzing the text, the
writing style, the scripture), chemistry and geology (providing details on the material upon which inscriptions were
carved), conservation (giving information about the state of the artefact), visualization and museology (about the
museums and places of conservation).

@@

\begin{figure}[!bp]
\centering
 \includegraphics[width=\columnwidth]{DamnjanovicetalEAGLE2016-img005.pdf}
\caption{Diagram of the metadata schema components and relationships.}
\label{fig:5}
\end{figure}

>>>>

%%%%%%%%
%TITLE:%
%%%%%%%%
\title{Integration of Multimedia Collections and Tools
for Interaction with Digital Content.
The case study of the Archaia Kypriaki Grammateia Digital Corpus}
\author[cyi]{Valentina Vassallo\corref{first}}
\author[cyi]{Uros Damnjanovic}
\author[cyi]{Sorin Hermon}
\date{2015-11-26}
\cortext[first]{Corresponding author. Email: (u.damnjanovic, v.vassallo, s.hermon@cyi.ac.cy}
\begin{document}

\maketitle
\begin{abstract}
Supporting a discovery, use, and navigation of digital collections is a fundamental part of providing access and
encouraging inquiry, interpretation, and knowledge. In this paper we present our efforts to store and explore
multimedia collections of archaeological data. Particularly, the case study of the Archaia Kypriaki Grammateia
epigraphic collection is presented. Our work can be seen twofold. One aspect of our work is to provide a place where
the data coming from various sources can be stored and accessed. Another aspect is to provide users with means to explore
this data. We argue that currently digital libraries are constrained by their webpage-based paradigm, thus not
providing the means for utilizing the full potential of the heritage data.
\end{abstract}
\keywords{Digital libraries, Data repository, Epigraphy, User interfaces, Data visualization, Interaction}

@@

\noindent Nowadays digital information format found its ways to all the spheres of our lives. Cultural heritage is no exception to
this. Gigabytes of data are being created on an every-day basis. For some time now, digital libraries are used to store
and provide access to digital heritage data. However, existing digital libraries often fail to provide the ability to
iteratively explore items, compare data trends, and engender the wisdom that comes from exploring data in new ways.
Current digital libraries are limited by their inadequate webpage-based paradigm, and it is easy for even the most
experienced scholar to get lost \citep{augmenting_2009}.
To overcome this issue, a new generation of digital
libraries should emerge that will serve as an interactive knowledge environment in which knowledge is created by
interacting with the data. Digital collections often lack features for deeper quantitative and qualitative analysis,
and even very useful functions, such as the ability to annotate or bookmark content, are often not supported. After
digitisation, these collections are typically monolithic, difficult to navigate, and can contain text which is of
variable quality in terms of language, spelling, punctuation and consistency of terminology.

According to \citet{fast_interaction_2010} digital libraries are seen as a store of epistemic potential with interaction having
the role to reveal much of the hidden complexity. In \citet{buchel_making_2014} the authors demonstrated the role that a
set of interactions can play in supporting users' understanding of spaces and their formation of cognitive maps when
working with map-based visualisations. According to \citet{algee_viewshare_2012} interactive visualizations empower users to
discover meaning and patterns within digital collections using dynamic, interactive displays. One example of utilizing
the interactive graphics in a real world digital library is presented in \citet{hienert_integrating_2012}. Authors developed
interactive visualization tools that support exploration of search queries and search results and help users in
formulating new queries. Another example, the application called DaisyViz shown in \citet{ren_daisyviz:_2010}, tries to help
users acquire better insights of the data by enabling users to rapidly develop domain-specific information
visualizations without traditional programming. In \citet{nasar_conceptual_2011} authors proposed a conceptual framework that
uses interactive visualizations for managing personal collections of images and videos with focus on data re-finding
and improved filtering. INVISQUE \citep{wong_invisque:_2011} is a novel system designed for interactive information
exploration. Instead of a conventional list-style arrangement, in INVISQUE information is represented by a
two-dimensional spatial canvas, with each dimension representing user-defined semantics. Search results are presented
as index cards, ordered in both dimensions. Intuitive interactions are used to perform tasks such as keyword searching,
results browsing, categorizing, and linking to online resources such as Google and Twitter. An approach for
overview-first exploration of data collections based on user-selected metadata properties is presented in \citet{bernard_content-based_2012}. In a 2D layout representing entities of the selected property are laid out based on their similarity with
respect to the underlying data content. The display is enhanced by compact summarizations of underlying data elements,
and forms the basis for exploratory navigation of users in the data space. In \citet{cao_whisper:_2012} authors propose a
novel visual design termed ``Whisper'' to fulfil the need for tracing information diffusion processes in social media, in
a real time manner. BirdVis \citep{ferreira_birdvis:_2011} leverages visualization techniques and uses them in a novel way to
better assist users in the exploration of interdependencies among model parameters. Furthermore, the system allows for
comparative visualization through coordinated views, providing an intuitive interface to identify relevant correlations
and patterns.

@@

\noindent STARC repository was initially developed to store the data coming from the daily work of the research group and as
content aggregator for various digital libraries projects: data from archaeological excavations, museums artefacts,
epigraphic corpora, etc. Data stored in the repository ranges from high resolution images, 3D models, texts, maps,
videos and audio resulting from various data acquisition procedures such as photogrammetry, laser scanning, 3D
modeling, photography, sketching, drawing and so on. \ Every item in the repository is described using a metadata
schema that depends on the type of data and according to the needs of the project. For most of the items, we are using
our own metadata schema (STARC metadata schemas) \citep{ronzino_metadata_2012, vassallo_revealing_2013}, that was developed to
thoroughly describe various aspects of data creation process. \ Once data is uploaded and available, it can also be
used in different systems. For example, the repository serves as source of data for aggregation procedures by which our
data feed the Europeana portal (\url{http://www.europeana.eu/portal/}) through different digital libraries projects (e.g.
Athena, CARARE, Linked Heritage, AthenaPlus, EAGLE). In order to facilitate the migration of data from heterogeneous
sources, the repository provides ingestion capabilities, that enable easy transfer of data from any available data
end-point. Another aspect of the back-end is the user management. It deals with providing different access levels for
different users’ groups. It also manages personal user space, where users can add their own information to the
repository and have access to it when exploring the data. Data annotations and personal collections are some of the
examples of how users can add information to the repository and use it for data exploration.

@@

\noindent We wanted to provide users with a number of tools that they can use while exploring the repository. In traditional web
based system, navigating from one tool to another usually means that we click on a link and then move to another page
where the tool is available. Once the user is on a page s/he can start interacting with the tool by adding or modifying
data or by performing any action relevant to the selected functionality. The process of exploring the repository starts
by running the initial query (Fig.~\ref{fig:1}).

\begin{figure}[!hbp]
\centering
 \includegraphics[width=\columnwidth]{DamnjanovicetalEAGLE2016-img001.png}
\caption{The exploration process starts by running the search. Once the results are available the user can add or remove
exploration tools. By default map tool is present in the view. The map tool is used to mark the geo-referenced data on
the map, and also to perform geo-search by defining the region of interest. }
\label{fig:1}
\end{figure}

Once the results are returned and displayed, the exploration process can start. By default the map functionality is
available in the view. When a new tool is added to the screen, it automatically gets an access to all the data
available in the current search. It also gets notified on any other tools that might be active in the view, so the
necessary communication channels are established. Once the results are ready, there are a number of filters available to
filter out the data, either by data type, collection. Search form can be used to filter out the data by typing the text
used to filter the data. As the user types, the data is automatically updated to show only those items that are related
to the text in the search field.

\begin{figure}[!hbp]
\centering
 \includegraphics[width=\columnwidth]{DamnjanovicetalEAGLE2016-img002.png}
\caption{The annotation tool. The user first selects the object in the search results, then selects the region of the
object and then adds the annotation. The annotation can be accessed when the annotated object is accessed, and by
specifying in the search that the search should also include annotations. }
\label{fig:2}
\end{figure}

The user can then run a new search repeatedly, and the search history functionality
will store all those searches for later reference. The map tool (Fig.~\ref{fig:1}) is used to show information about the
geo-referenced data. It provides two main functions. One is that it accompanies the search tool. Whenever the new
search is ran, and the results become available, the map tool updates itself with new data, showing the geographical distribution of the search results. The user can then use the map to filter out the data by clicking on the
markers on the map. Also the map can also be used as a search tool. By selecting the search map mode, the rectangle
mask appears on the map, that is used to define the region of interest on the map. This region is then used as an
additional parameter for the search.

@@

@@

\begin{figure}[!hbp]
\centering
 \includegraphics[width=\columnwidth]{DamnjanovicetalEAGLE2016-img003.png}
\caption{Visualization tool. Visualization tool is used to accompany the search tool by showing live visualizations as
the search results change. When the search is ran and the search results are available the visualization tools
automatically update itself and shows the visualisation related to the current search results. }
\label{fig:3}
\end{figure}

@@

\begin{figure}[!hbp]
\centering
 \includegraphics[width=\columnwidth]{DamnjanovicetalEAGLE2016-img004.jpg}
\caption{Multimedia data constituting the AKG digital collection stored in the STARC repository.}
\label{fig:4}
\end{figure}

@@

The metadata schema is organized in groups corresponding to different research areas clustered in wrappers and
sub-wrappers, in order to fully describe all information that the ‘asset inscription’ contains. This schema,
distinguished by its multidisciplinary structure, include all disciplines relevant for description and representation
of epigraphies: archaeology (investigating for example the context of the finds), philology s(analyzing the text, the
writing style, the scripture), chemistry and geology (providing details on the material upon which inscriptions were
carved), conservation (giving information about the state of the artefact), visualization and museology (about the
museums and places of conservation).

@@

\begin{figure}[!hbp]
\centering
 \includegraphics[width=\columnwidth]{DamnjanovicetalEAGLE2016-img005.pdf}
\caption{Diagram of the metadata schema components and relationships.}
\label{fig:5}
\end{figure}

### 2016-01-19 source/8/EAGLE_2016_submission_8_revised.tex
papers revised
minor fixes

% dummy text (remove in a normal thesis)
% remove if not necessary
\usepackage{siunitx}
%Natbib for bibliography management
\usepackage[authoryear]{natbib}
% custom commands
\newcommand{\bs}{\textbackslash}

@@

%all title info missing!
\title{A VIRTUAL RESEARCH ENVIRONMENT TO DOCUMENT AND ANALYZE NON-ALPHABETIC WRITING SYSTEMS. A Case Study for Maya Writing.}
\author[bonn]{Katja Diederichs\corref{first}}
\author[bonn]{Sven Gronemeyer}
\author[bonn]{Christian Prager}
\author[bonn]{Elisabeth Wagner}
\author[bonn]{Franziska Diehr}
\author[bonn]{Brodhun Maximilian}
\author[bonn]{Nikolai Grube}

\address[bonn]{missing}

\cortext[bonn]{missing}

\begin{document}

\maketitle

@@

\begin{figure}[!bp]
\centering
 \includegraphics[width=\columnwidth]{EAGLE2016submission8revisedx-img001.png}
\caption{Visualization of the important steps in the workflow of epigraphic analysis, drawings by \citet{Prager2015}}
\label{fig:1}
\end{figure}

\subsection[Conceptual Questions]{Conceptual Questions}
When developing and technically implementing the digital epigraphic research environment, the major focus is on pursuing
the following objectives: 
\begin{enumerate}
\item creation of a markup schema to represent these steps (which in practice are interconnected
and may also occur parallel to each other in the form of alternative interpretations); 
\item defining of the precise requirements that a data structure must fulfill in order to sufficiently represent complex
i.e. non-alphabetic textual data for epigraphic analysis and research. 
\end{enumerate}

@@

\subsubsection[1. Allography of Graphemes ]{1. Allography of Graphemes }
% THESE ARE SUBSUB*SUB*SECTIONS...
A very important example of a subject of graphemic research is provided by allography, a phenomenon which describes a
1:n relationship between a grapheme and its various graph representations \citep{Crystal1997}.

@@

\subsubsection[2. Graphotactic Strategies of a Writing System ]{2. Graphotactic Strategies of a Writing System
}
For analysis, it is important to convey the graphemics of a writing system, which are lost in an alphabetic
transliteration. Graphemics concerns sign function, or graphotactic strategies for constituting a meaningful unit or
word. Representing the reading order of signs in the text and in each block is similarly imperative. To
graphotactically link individual signs in a hieroglyphic block in Maya writing, various representational rules may be
used (see Fig.~\ref{fig:2}), e.g. affixation, infixation, conflation, or superimposition \citep{Zender1999}. Additional typical,
functional traits of logo-syllabic writing systems include underspellings, diacritics, phonemic complements and
indicators, semantic classifiers and determinatives, and sign class convergences. However, these characteristics may
have very diverse graphotactical manifestations in different writing systems, which nonetheless need to be standardized
and communicated in a digital structure in order that researchers may thus take note of various interpretations under
discussion.

\begin{figure}[!bp]
\centering
 \includegraphics[width=\columnwidth]{EAGLE2016submission8revisedx-img002.png}
\caption[Some possible sign combinations in Classical Mayan hieroglyphs after \citet{Zender1999}, drawings by
\citet{Prager2015}]{Fig. 1.2. Some possible sign combinations in Classical Mayan hieroglyphs after \citet{Zender1999}, drawings by \citet{Prager2015}}
\label{fig:2}
\end{figure}

>>>>

%Natbib for bibliography management
\usepackage[authoryear]{natbib}
% custom commands
\newcommand{\bs}{\textbackslash}

@@

%all title info missing!
\title{A Virtual Research Environment to 
Document and Analyze Non-alphabetic Writing Systems. 
A Case Study for Maya Writing}
\author[BONN]{Katja Diederichs\corref{first}}
\author[BONN,MEL]{Sven Gronemeyer}
\author[BONN]{Christian Prager}
\author[BONN]{Elisabeth Wagner}
\author[GOT]{Franziska Diehr}
\author[GOT]{Maximilian Brodhun}
\author[BONN]{Nikolai Grube}
%
\address[GOT]{Niedersächsische Staats- und Universitätsbibliothek Göttingen}
\address[MEL]{La Trobe University Melbourne}
\address[BONN]{Rheinische Friedrich-Wilhelms-Universität Bonn}
%
\cortext[first]{Corresponding author. Email: diederichs@uni-bonn.de}

\begin{document}
\maketitle

@@

\begin{figure}[!hbp]
\centering
 \includegraphics[width=\columnwidth]{EAGLE2016submission8revisedx-img001.png}
\caption{Visualization of the important steps in the workflow of epigraphic analysis, drawings by \citet{Prager2015}}
\label{fig:1}
\end{figure}

\subsection[Conceptual Questions]{Conceptual Questions}
When developing and technically implementing the digital epigraphic research environment, the major focus is on pursuing
the following objectives: 
\begin{enumerate}
\item Creation of a markup schema to represent these steps (which in practice are interconnected
and may also occur parallel to each other in the form of alternative interpretations); 
\item Defining of the precise requirements that a data structure must fulfill in order to sufficiently represent complex
i.e. non-alphabetic textual data for epigraphic analysis and research. 
\end{enumerate}

@@

\vspace{0.5cm}
\noindent\textbf{Allography of Graphemes }\\
% THESE ARE SUBSUB*SUB*SECTIONS...
A very important example of a subject of graphemic research is provided by allography, a phenomenon which describes a
1:n relationship between a grapheme and its various graph representations \citep{Crystal1997}.

@@

\vspace{0.5cm}
\noindent\textbf{Graphotactic Strategies of a Writing System}\\
For analysis, it is important to convey the graphemics of a writing system, which are lost in an alphabetic
transliteration. Graphemics concerns sign function, or graphotactic strategies for constituting a meaningful unit or
word. Representing the reading order of signs in the text and in each block is similarly imperative. To
graphotactically link individual signs in a hieroglyphic block in Maya writing, various representational rules may be
used (see Fig.~\ref{fig:2}), e.g. affixation, infixation, conflation, or superimposition \citep{Zender1999}. Additional typical,
functional traits of logo-syllabic writing systems include underspellings, diacritics, phonemic complements and
indicators, semantic classifiers and determinatives, and sign class convergences. However, these characteristics may
have very diverse graphotactical manifestations in different writing systems, which nonetheless need to be standardized
and communicated in a digital structure in order that researchers may thus take note of various interpretations under
discussion.

\begin{figure}[!hbp]
\centering
 \includegraphics[width=\columnwidth]{EAGLE2016submission8revisedx-img002.png}
\caption[Some possible sign combinations in Classical Mayan hieroglyphs after \citet{Zender1999}, drawings by
\citet{Prager2015}]{Fig. 1.2. Some possible sign combinations in Classical Mayan hieroglyphs after \citet{Zender1999}, drawings by \citet{Prager2015}}
\label{fig:2}
\end{figure}

### 2016-01-19 source/9/Latin epigraphy for the visually impaired.tex
more biblio added, papers compiled with bibliography
papers revised
minor fixes

\section{Introduction}
\noindent The Declaration of the Rights of Deaf-Blind Persons, adopted by the UN in 1979, provides that every blind or deaf person
has the right to enjoy the same privileges, guaranteed to all people and to have their aspirations and abilities
recognized and respected. These principles are reaffirmed by the UN Convention on the Rights of Persons with
Disabilities of 2006, which states (Art. 30) that the Member States must recognize ``the right of persons with
disabilities to take part on an equal basis with others in cultural life and should take all appropriate measures to
ensure to persons with disabilities:

@@

\item[C] enjoy access to places of cultural activities, such as theaters, museums, cinemas, libraries and tourism services,
and, as far as possible, to monuments and sites of national cultural importance.''

@@

The problem of accessibility and enjoyment of works of art by the visually impaired has long existed. Sculptures, which
are sometimes designed to be touched, and sometimes are permitted to be touched, provide the possibility for blind
viewers to benefit directly from the experience of the original artwork. For two-dimensional paintings this is of
course not possible. It is necessary to create three-dimensional resin reliefs. This same issue has recently been
understood to exist also in other artistic disciplines, including that of archaeology. This of course includes
epigraphy, which is the study of all those materials which have written inscriptions. Most relevant to this is Latin
Epigraphy, simply because it is the one closest to our language and which uses the same graphic signs of our own
writing.

@@

The best solution to make the epigraphic media available to the visually impaired turns out to be, therefore, one that
involves more visitors who can better understand the artwork. In this regard, it being understood that the chance of
touching an original work wherever possible is a uniquely fascinating act, it becomes necessary to propose
reproductions that can make the material accessible by following the concept of ``design for all,'' which is adaptable to
all types of art. As rightly noted by the scholar and educator Enzo Tioli \citep{Tioli2006},
this principle can be defined as ``identical, whenever possible, equivalent when that is not possible.'' All that can be
touched and seen in the original work constitutes the ``identical'', the use of a copy constitutes the ``equivalent''.

\section{The Techniques of Accessibility}
\noindent In the world of epigraphy, casts of inscriptions are often made use of, both to facilitate their study by specialists,
and also to replace those exhibits that had been located outside and brought indoors to protect them from weather
conditions. These casts, if used by the visually impaired, present the same problems of accessibility as the original
in terms of the difficulty of understanding the incision’s grooves. It is necessary, therefore, to have a reproduction
of a different type, upon which the letters can be transliterated in relief.

@@

This technique would be very valid when used in the case of reproduction of mosaics and inscriptions in the mosaic.
These reproductions make for the best situation where the user can differentiate between the various levels. The
background can be made on one level, while the drawn design can be made on another and also possible written upon. The
use of 3D would obtain a result similar to that by the Federazione Nazionale delle Istituzioni Pro Ciechi (National
Federation of Institutions for the Blind) NPO in the Square of the Corporations of Ostia Antica. In this area, paved
with many squares of mosaics, which bear witness to the intense commercial activity of ancient Ostia, the reproduction
of Ostia’s famous lighthouse was created with different levels of elevation of a mosaic in black and white, which allow
the user to not only to understand the image, but also to understand the technique of its realization. In this case,
though, the reproduction does not contain writing, because the chosen image contained no inscriptions. But nothing
prevents the use of the same technique using written entries, notwithstanding the need for an additional explanatory
panel. The disadvantage of this technique is the high cost of implementation, given that it is a reproduction created
exclusively from a specific work, and requiring a rather high level of craftsmanship since it is not merely modeled
from a metal matrix.

@@

\nocite{accessibilità,Bellini2000,Cetorelli2004,Hersh2008,Jansson2003,Levi2015,Velestino2007}

>>>>

\section{Introduction}
\noindent The Declaration of the Rights of Deaf-Blind Persons, adopted by the UN in 1979, provides that every blind or deaf person
has the right to enjoy the same privileges, guaranteed to all people and to have their aspirations and abilities
recognized and respected. These principles are reaffirmed by the UN Convention on the Rights of Persons with
Disabilities of 2006, which states (Art. 30) that the Member States must recognize

\begin{quotation}
the right of persons with disabilities to take part on an equal basis with others in cultural life and should take all appropriate measures to
ensure to persons with disabilities:

@@

\item[C] enjoy access to places of cultural activities, such as theaters, museums, cinemas, libraries and tourism services,
and, as far as possible, to monuments and sites of national cultural importance.

@@

\end{quotation}
The problem of accessibility and enjoyment of works of art by the visually impaired has long existed. Sculptures, which
are sometimes designed to be touched, and sometimes are permitted to be touched, provide the possibility for blind
viewers to benefit directly from the experience of the original artwork. For two-dimensional paintings this is of
course not possible. It is necessary to create three-dimensional resin reliefs. This same issue has recently been
understood to exist also in other artistic disciplines, including that of archaeology. This of course includes
epigraphy, which is the study of all those materials which have written inscriptions. Most relevant to this is Latin
Epigraphy, simply because it is the one closest to our language and which uses the same graphic signs of our own
writing.

@@

The best solution to make the epigraphic media available to the visually impaired turns out to be, therefore, one that
involves more visitors who can better understand the artwork. In this regard, understood that the chance of
touching an original work wherever possible is a uniquely fascinating act, it becomes necessary to propose
reproductions that can make the material accessible by following the concept of ``design for all,'' which is adaptable to
all types of art. As rightly noted by the scholar and educator Enzo Tioli \citep{Tioli2006},
this principle can be defined as ``identical, whenever possible, equivalent when that is not possible.'' All that can be
touched and seen in the original work constitutes the ``identical'', the use of a copy constitutes the ``equivalent''.

\section{The Techniques of Accessibility}
\noindent In the world of epigraphy, casts of inscriptions are often used, both to facilitate their study by specialists,
and also to replace those exhibits that had been located outside and brought indoors to protect them from weather
conditions. These casts, if used by the visually impaired, present the same problems of accessibility as the original
in terms of the difficulty of understanding the incision’s grooves. It is necessary, therefore, to have a reproduction
of a different type, upon which the letters can be transliterated in relief.

@@

This technique would be very valid when used in the case of reproduction of mosaics and inscriptions in the mosaic.
These reproductions make for the best situation where the user can differentiate between the various levels. The
background can be made on one level, while the drawn design can be made on another upon which it is possible to write. The
use of 3D would obtain a result similar to that by the Federazione Nazionale delle Istituzioni Pro Ciechi (National
Federation of Institutions for the Blind) NPO in the Square of the Corporations of Ostia Antica. In this area, paved
with many squares of mosaics, which bear witness to the intense commercial activity of ancient Ostia, the reproduction
of Ostia’s famous lighthouse was created with different levels of elevation of a mosaic in black and white, which allow
the user not only to understand the image, but also to understand the technique of its realization. In this case,
though, the reproduction does not contain writing, because the chosen image contained no inscriptions. But nothing
prevents the use of the same technique using written entries, notwithstanding the need for an additional explanatory
panel. The disadvantage of this technique is the high cost of implementation, given that it is a reproduction created
exclusively from a specific work, and requiring a rather high level of craftsmanship since it is not merely modeled
from a metal matrix.

@@

\nocite{accessibilita,Bellini2000,Cetorelli2004,Hersh2008,Jansson2003,Levi2015,Velestino2007}
